Exploring probability distributions for bivariate temporal granularities

### 2019-10-30 main.Rmd
harmony and clash section update

>>>>

When we plot some observations with two circular or aperiodic granularities used as aesthetics or facets, problems arise when we have empty combinations. It is important to check how these two granularities interact with each other, before we expect to gain insights by plotting them together.

For example, suppose we have two circular or aperiodic granularities $C_1$ and $C_2$, such that $C_1$ maps row numbers to a set $\{A,B,C,D$\}, and $C_2$ maps row numbers to a set $\{X,Y,Z\}$.  That is, let $S_{ij}$ be the set of row numbers such that for all $s \in S_{ij}$, $C_1(s) = i$ and $C_2(s) = j$. In this example, we have 12 such sets $S_{ij}$ because $i$ can take 4 values and $j$ can take 3 values. The graphs that don't work are those where many of these 12 sets are empty. Now, many situations can lead to any of these sets being empty. Let us discuss the following cases, where one or more of these sets can be empty.

Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
These are called structurally empty combinations. Let us take a specific example, where $C_1$ maps row numbers to Day-of-Month and $C_2$ maps row numbers to Week-of-Month. Here $C_1$ can take 31 values while $C_2$ can take 5 values. There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of WOM and DOM. Many of these are empty. For example $S_{1,5}$, $S_{21,2}$, etc. In fact, most of these 155 sets will be empty, making the combination of $C_1$ and $C_2$ in a graph unhelpful. These are structurally empty sets in that it is impossible for them to have any observations.

Secondly, empty combinations can turn up due to difference in occurance of event location or duration in a calendar. These are called event-driven empty combinations. Again, let us consider a specific example to illustrate this. Let $C_1$ be DOW and $C_2$ be WorkingDay/NonWorkingDay. Here $C_1$ can take 7 values while $C_2$ can take 2 values. So there are 14 sets $S_{ij}$ corresponding to the possible combinations of DOW and WD/NWD. While potentially all of these can be non-empty (it is possible to have a public holiday on any DOW), in practice many of these combinations will probably have very few observations. For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne.

<!-- However, this is different from the first example because these are not structurally empty. Also, the graphic may still be useful if we want to look at whether the interaction of $C_1$ and $C_2$ affects the distribution we are plotting. -->

Thirdly, empty combinations can be a result of how granularities are constructed. Let $C_1$ maps row numbers to "Business days", which are days from Monday to Friday except holidays and $C_2$ is Day-of-Month. Then the weekends in Days-of-Month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed. this is different from the first case because these are not structurally empty combinations. The structure of the calendar does not lead to these missing combinations, but the construction of the granularity does. Hence, they are referred to as "build-based" empty combinations.

Another example could be where $C_1$ maps row numbers to Day-of-Week and $C_2$ maps row numbers to Month-of-Year. Here $C_1$ can take 7 values while $C_2$ can take 12 values. So there are $12\times7=84$ sets $S_{ij}$ corresponding to the possible combinations of DOW and MOY. All of these are non-empty because every DOW can occur in every month. So graphics involving $C_1$ and $C_2$ are potentially useful.
